---
title: "Reproducable Research"
author: "Mathijs Haerden"
date: "4 juli 2018"
output:
  html_document: default
  pdf_document: default
---

Load required packages:
```{r echo=TRUE, message=FALSE}
library(ggplot2)
library(plyr)
```


Reading in the data and transforming the date into a date value
```{r echo=TRUE}
data <- read.csv("activity.csv")
data$date <- as.Date(as.character(data$date))
```

The number of steps taken each day is as follows:
```{r echo=TRUE}
totalPerDay <-tapply(data$steps, data$date, sum)
df<- data.frame(rownames(totalPerDay), totalPerDay)
names(df) <- c("Date", "Total Steps")
rownames(df) <- NULL
print(df)
```

Now put it in a histogram
```{r echo=TRUE}
hist(df$"Total Steps", main ="Histogram of total steps taken per day", xlab = "Total Steps")
```

Report the mean and median of the total steps taken per day:
```{r echo=TRUE}
mean(df$'Total Steps', na.rm = TRUE)
median(df$'Total Steps', na.rm = TRUE)
```

Create a time series plot of the 5-minute interval and the average steps taken
```{r echo=TRUE}
data2 <- ddply(data, ~ interval, summarise, avg=mean(steps, na.rm=TRUE)) 
plot(data2$interval, data2$avg, type = "l", xlab="5-minute interval", ylab = "Average steps taken", main = "Time series plot of 5-minute interval and average steps taken")
```

The 5-minute interval with on average the maximum steps taken is:
```{r echo=TRUE}
as.numeric(data2[data2$avg==max(data2$avg),1])
```

Count the total number of values in the dataset:

```{r echo=TRUE}
table <- table(is.na(data$steps))
table <- table["TRUE"]
names(table) <- "Number of NA"
print(table)
```

Replace missing values with average per 5-minute interval. A new dataset is created and a table shows that there are no missing values left
```{r echo=TRUE, message=FALSE}
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data3 <- ddply(data, ~ interval, transform, steps = impute.mean(steps))
table(is.na(data3))
```

A historgram on the new dataset shows that the data has changed
```{r echo=TRUE}
totalPerDay <-tapply(data3$steps, data3$date, sum)
df<- data.frame(rownames(totalPerDay), totalPerDay)
names(df) <- c("Date", "Total Steps")
rownames(df) <- NULL

hist(df$"Total Steps", main ="Histogram of total steps taken per day", xlab = "Total Steps")
```

The mean has not changed, the median slightly
```{r echo=TRUE}
mean(df$'Total Steps')
median(df$'Total Steps')
```

Setting the locale for English weekdays
```{r echo=TRUE, results='hide'}
Sys.setlocale("LC_TIME", "C")
```


Create new variable which indicates whether its a weekday or a weekend day
```{r echo=TRUE}
data3$daytype <- weekdays(data3$date)
data3$daytype <- ifelse(data3$daytype == "Saturday" | data3$daytype =="Sunday", "Weekend", "Weekday")
```

Create panel plot
```{r echo=TRUE}
data4 <- ddply(data3, .(interval,daytype), summarise, avg=mean(steps, na.rm=TRUE)) 
ggplot(data4, aes(x=interval, y=avg)) + geom_line()+ facet_grid(daytype~.)
```



